%% -*- fill-column: 120; -*-
\section{Introduction}
\label{sec_introduction}

Over the last several years, cryptographic currencies like Bitcoin~\cite{} started a disruption that has grown
well-beyond currency. Inspired by the decentralization of fiduciary authority represented by Bitcoin, businesses and
governments across the world are investigating disintermediated and decentralized services. Usages are diverse:
federated identity, clearing and settlement of financial transactions, multi-party supply chain management, provenance
of food, and privacy-preserving access to medical records. At the core of this revolution is a data structure called a
{\it blockchain}. Simply put, a blockchain is a data structure that uses cryptographic hashes to create an immutable
ordering of events. Communities use blockchains to represent a shared view of ``truth''. For example, when events like
``transfer $50 from john to jane'' or ``grant tom access to file A'' are committed to a blockchain, the entire community
agrees that the event occurred and can act accordingly.

There are many ways for a community to build and maintain a blockchain. In some cases, a trusted organization (like a
representative consortium or a contracted infrastructure provider) can build a blockchain using fairly traditional
distributed database technologies. However, for those cases where no trusted organization exists building a blockchain
requires a protocol that creates consensus across the community about the current, correct state of the blockchain.
There are many different protocols for building consensus based on requirements related to performance, scalability,
consistency, threat model, and failure model. Traditionally, protocols that achieve consensus for arbitrary
(i.e. Byzantine) faults require several rounds of explicit voting. While this approach provides high throughput and
finality of commits, its reliance on a well-known, generally static voting block and the heavy use of communication
(multiple rounds of N^2 messages) limit its use to reletively small, localized, and stable applications. Bitcoin
introduced a new method for achieving consensus, called Nakamoto consensus, that supports broad participation from a
dynamic group of participants. Nakamoto consensus relies on uncoordinated, distributed leader election that is efficient
for communication, but often very computationally intense. For example, Bitcoin uses ``proof of work'', a computation to
solve a cryptographic puzzle through repetitive ``guess and test''; a process that consumes gigawatts of power. The
elected leader is granted unilateral authority to propose the next block in the blockchain. 

This paper describes a new leader election algorithm designed for Nakamoto style consensus, called ``Proof of Elapsed
Time'' or PoET, that elects leaders through a simple computation in a trusted execution environment provided by
commodity processors. PoET preserves the fairness and integrity of leader election provided by proof-of-work without
high cost of electricity and custom hardware. This allows deployment of highly-resilient, fault tolerant applications
for a dynamic community using commonly available computing resources. 

% -----------------------------------------------------------------
Unfortunately, there are a variety of limitations that handicap the block chain in its current
instantiation and prevent it from being used for other (potentially) useful purposes. We focus on
leveraging Intel's SGX environment [AGJS13] to re-engineer the proof of work to a proof of wait,
creating a new lottery-based consensus protocol. This will provide greater efficiency and reduce
the amount of energy used and high costs that we currently see.

The key concept of the proof of wait is a guaranteed wait time, provided through the Trusted
Execution Environment (TEE) that we get with SGX. Each \miner" will request a wait time from
the TEE, and the miner with the lowest wait time \wins," thus getting to create the next block.
In order to prove that the node did indeed wait the claimed time, an attestation is also created
within the TEE. This validation will be done using the Enhanced Privacy ID (EPID) [BL07, BL09]
scheme


Consensus is the process of building agreement amongst a group of mutually distrusting participants. 
% say something about blockchain applications

There are many different protocols for building consensus based on requirements related to performance, scalability,
consistency, threat model, and failure model. Traditionally, protocols that achieve consensus for arbitrary
(i.e. Byzantine) faults require several rounds of explicit voting. While this approach provides high throughput and
finality of commits, its reliance on a well-known, generally static voting bloc and the heavy use of communication
(multiple rounds of N^2 messages) limit its use to reletively small, localized, and stable applications. Bitcoin
introduced a new method for achieving consensus, called Nakamoto consensus, that supports broad participation from a
very dynamic group of participants. Nakamoto consensus relies on uncoordinated, distributed leader election that is
often very computationally intense. For example, Bitcoin uses ``proof of work'', a computation to solve a cryptographic
puzzle through repetitive ``guess and test''; a process that consumes gigawatts of power.

This paper describes a new leader election algorithm designed for Nakamoto style consensus, called ``Proof of Elapsed
Time'' or PoET, that elects leaders through a simple computation in a trusted execution environment provided by
commodity processors. PoET preserves the fairness and integrity of leader election provided by proof-of-work without
high cost of electricity and custom hardware. This allows deployment of highly-resilient, fault tolerant applications
for a dynamic community using commonly available computing resources. 



Bitcoin introduced an alternative approach, often called Nakamoto consensus~\cite{}, where participants implicitly vote
by building on a particular chain. 
%Open participation, O(N) messages with high probability, 

often referred to as Nakamoto consensus, elects
a leader through some form of ``lottery''. The leader then proposes a block that can be added to a chain of previously
committed blocks.


Nakamoto consensus was introduced in Bitcoin, where the first participant to successfully solve a cryptographic
puzzle wins the leader-election lottery. The elected leader broadcasts the new block to the rest of the participants who
implicitly vote to accept the block by adding the block to a chain of accepted blocks and proposing subsequent
transaction blocks that build on that chain.

The first, called PoET for ``proof of elapsed time'', is a lottery protocol that builds on trusted execution
environments provided by Intel's Software Guard Extensions (SGX) to address the needs of large populations of
participants.

Sawtooth Lake provides a Nakamoto consensus algorithm called PoET. As with Bitcoin PoET uses a lottery for leader
election based on a guaranteed wait time provided through a trusted execution environment. 

For the purpose of achieving distributed consensus efficiently, a good lottery function has several characteristics:

\begin{itemize}
\item Fairness:
  The function should distribute leader election across the broadest possible population of participants.

\item Investment:
  The cost of controlling the leader election process should be proportional to the value gained from it. 

\item Verification:
  It should be relatively simple for all participants to verify that the leader was legitimately selected. 
\end{itemize}

PoET uses a trusted execution environment (the SLDL implementation uses Intel's SGX) to ensure the safety and randomness
of the leader election process without requiring the costly investment of power and specialized hardware inherent in
most ``proof'' algorithms.

Every PoET validator requests from a trusted function a random time to wait before claiming leadership. The validator
with the shortest wait time for a particular transaction block is implicitly elected the leader. Figure 4 provides
pseudo-code for two functions that could be executed within a trusted execution environment. The function
``CreateTimer'' creates a timer for a transaction block that is guaranteed to have been created by the TEE. The
``CheckTimer'' function verifies that the timer was created by the TEE and, if it has expired, creates an attestation
that can be used to verify that validator did, in fact, wait the allotted time before claiming the leadership role.

The PoET leader election algorithm meets the criteria for a good lottery algorithm. It randomly distributes leadership
election across the entire population of validators with distribution that is similar to what is provided by other
lottery algorithms. The probability of election is proportional to the resources contributed (in this case, resources
are general purpose processors with a trusted execution environment). An attestation of execution provides information
for verifying that the certificate was created within the TEE (and that the validator waited the allotted
time). Further, the low cost of participation increases the likelihood that the population of validators will be large,
increasing the robustness of the consensus algorithm.
